---
title: "Russian River sockeye run timing 2018"
author: "Andy Barclay"
date: "January 23, 2019"
output: html_document
---
This R Markdown documents the genetic mixed stock analysis of sockeye salmon sampled from the lower Kenai River test fishery in 2018.  The purpose of this study is to determine if non-Russian River sockeye are present in the Kenai River during the early run.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

##Set up workspace
###Get functions and load tidy packages
```{r workspace setup}

source("C:\\Users\\awbarclay\\Documents\\R\\GitHubCloneFunctions.R")#GCL functions
source("V:\\Analysis\\Staff\\Andy Barclay\\R\\New Functions\\WorkspaceFolders.GCL.R")#A function I wrote

library("tidyverse")

```

###Create output folders 
```{r create folders, echo=TRUE, message=FALSE}

#WorkspaceFolders.GCL(Folders=c("data",output","bayes","objects","rubias"),Subfolders=list(bayes=c("baseline","mixture","control","output"),rubias=c("baseline","mixture","output")),wd=getwd())

```
###Create tissue import file with CAPTURE_DATE for SKENNT18L
```{r read in original tissue table}

tissue_table_original <- read_csv(file="data/SKENNT18L_fish_and_tissues_report_1.23.19.csv")

header<-dimnames(tissue_table_original)[[2]]

```
```{r read in data sheet}

data_sheet <- read_csv(file="data/2018 Kenai Sockeye Netting-Lower selection.csv")

```
```{r add capture date}

tissue_table_new <- tissue_table_original %>% 
  left_join(y=data_sheet,by=c("DNA_TRAY_CODE","DNA_TRAY_WELL_CODE")) %>% 
  mutate(CAPTURE_DATE=CAPTURE_DATE.y,CAPTURE_LOCATION="Lower_Kenai_River_Test_Fish") %>% 
  select(header)

write_excel_csv(tissue_table_new,"data/SKENNT18L_tissue_table_import_1.23.19.csv",na="")

```


###Create locus control and get genotype data
```{r locus control, echo=TRUE}

#CreateLocusControl.GCL(markersuite = "Sockeye2011_96SNPs", username ="awbarclay", password = password)#Locus control

#LOKI2R.GCL(sillyvec="SKENNT18L",username="awbarclay",password)#Pull Data from LOKI

#save.image("V:/Analysis/2_Central/Sockeye/Cook Inlet/2012 Baseline/Mixture/Russian_River_sockeye_run_timing_2018/Russian_River_sockeye_run_timing_2018.Rdata")

load("V:/Analysis/2_Central/Sockeye/Cook Inlet/2012 Baseline/Mixture/Russian_River_sockeye_run_timing_2018/Russian_River_sockeye_run_timing_2018.Rdata")

```

###Create initial locus object
```{r initial locus object}

loci <- LocusControl$locusnames

```

##Data cleanup
###Check initial sample size
```{r initial sample size, echo=TRUE}

ColSize<-data.frame(row.names = "SKENNT18L")
ColSize$Original_N <- SKENNT18L.gcl$n
ColSize

```

###Removing individuals with <80% of loci with gentotypes
Fish that have gentoypes for less than 80% of loci are likely to have poor quality DNA and might introduce gentotyping errors into the mixture data and reduce the accruacy of the mixed stock analysis (MSA)
```{r missloci,echo=TRUE}

MissLOCI=RemoveIndMissLoci.GCL(sillyvec="SKENNT18L",proportion=0.8)
MissLOCI

ColSize$AfterMissLoci_N <- SKENNT18L.gcl$n
ColSize

```

###Check for duplicate individuals and remove them
Fish with 99% of scores that match
No duplicates to remove
```{r dupckeck, echo=TRUE, message=FALSE}

dupcheck<-CheckDupWithinSilly.GCL(sillyvec="SKENNT18L",loci=loci,quantile=NULL,minproportion=0.99)
dupcheck
```
```{r final sample size}
ColSize$Final_N <- SKENNT18L.gcl$n
ColSize

```
##Combine MHCs and mtDNA loci
```{r}

CombineLoci.GCL(sillyvec = "SKENNT18L",markerset = loci[c(10,12,13)],update=TRUE)#mtDNA

CombineLoci.GCL(sillyvec = "SKENNT18L",markerset = loci[c(32,33)],update=TRUE)#mtDNA

loci93 <- LocusControl$locusnames[-c(10,12,13,32,33)]

```


##Create BAYES baseline and get objects from baseline workspace
```{r get objects from baseline }

attach("V:/Analysis/2_Central/Sockeye/Cook Inlet/2012 Baseline/Baseline/EarlyRussianProoftests.RData")

groups<-groups

groupvec <- groupvec

Kenai14pops<-Kenai14pops

PopNames <- LocNames

basefortran<-CreateBaseline.GCL(sillyvec=Kenai14pops,loci=loci93,dir="bayes/baseline",basename="Kenai14pops96loci",type="BAYES",groupvec=NULL)#BAYES
baseline<-create_rubias_baseline(sillyvec = Kenai14pops,loci=loci93,group_names = groups,groupvec = groupvec,baseline_name = "Kenai14pops96loci")#rubias

detach()
baseline
```
##Set up mixture objects
###Pool to mixture .gcl objects
```{r mixture setup}

require(lubridate)

attr <- as.tibble(SKENNT18L.gcl$attributes) %>% 
  mutate(CAPTURE_DATE=as.Date(CAPTURE_DATE))

Mixtures=c("May24toJun3","Jun4to10","June11to20","Jun21to30")

mixture_dates <- data_sheet %>% 
  group_by(STRATUM) %>% 
  summarize(Start_Date=min(CAPTURE_DATE),End_Date=max(CAPTURE_DATE)) %>% 
  mutate(Mixture=Mixtures)

for(mix in Mixtures){

  dates <- mixture_dates %>% 
    filter(Mixture==mix)
  
  PoolCollections.GCL(collections="SKENNT18L",
                      loci=loci93,
                      IDs=attr %>% 
                        filter(CAPTURE_DATE%in%with(dates, seq(from=Start_Date,to=End_Date,by=1))) %>% 
                        pull(FK_FISH_ID) %>%
                        as.character%>% 
                        list,
                      newname = mix
                      )
}

sapply(Mixtures,function(mix){get(paste0(mix,".gcl"))$n})#Mixture sample sizes

```
###Create mixture inputs
####BAYES mixture files
```{r create BAYES mixture input files,echo=FALSE}

for(mix in Mixtures){
  
  mixfortran<-CreateMixture.GCL(sillys = mix,loci=loci93,mixname=mix,dir="bayes/mixture")
  
}

```
####rubias mixture input files and object
```{r create rubias mixture input files,echo=FALSE}

mix.in<-create_rubias_mixture(sillyvec=Mixtures,loci=loci93,path = "rubias/mixture")
mix.in

```

##Create BAYES Control Files
```{r bayes control files,echo=FALSE}

initmat<-MultiChainInits.GCL(npops=length(Kenai14pops),nchains=5,prop=.9)#Initial start values

priors<-Prior.GCL(groupvec=groupvec,groupweights=rep(1/max(groupvec),max(groupvec)),minval=0.01)#Flat prior


for(mix in Mixtures){
  
 
  CreateControlFile.GCL (sillyvec=Kenai14pops,loci=loci93,mixname=mix,basename="Kenai14pops96loci",suffix="",nreps=40000,nchains=5,groupvec=groupvec,priorvec=priors,initmat=initmat,dir="bayes/control",seeds=matrix(sample(seq(10000),3*5),nrow=3),
                               thin=c(1,1,1),mixfortran=mixfortran,basefortran=basefortran,switches="F T F T T T F")
  
}

```
###Move BAYES file to output folder for analysis
```{r move BAYES files}
#Move files into output folders for running BAYES

 folders<-c("control","mixture")
 dir = paste0(getwd(),"/bayes")
 
 bayesexe.paths=list.files("V:\\Analysis\\2_Central\\Chinook\\Cook Inlet\\2014\\Baseline\\BAYES\\BAYES exe files",full.names=T,include.dirs=T)
 
 bayesexe.files=list.files("V:\\Analysis\\2_Central\\Chinook\\Cook Inlet\\2014\\Baseline\\BAYES\\BAYES exe files",full.names=F,include.dirs=T)


#Move mixture and control files to output folder
 for(mix in Mixtures){
   
   dir.create(path=paste0(dir,"/output/",mix))
 
    for(folder in folders){
          
          files=list.files(paste(dir,"/",folder,sep=''),pattern=mix) 
          
          paths=paste(dir,"/",folder,"/",files,sep='')
          
          for (path in paths){
            
            file.rename(from= path,to=sub(pattern=paste("bayes/",folder,sep=''),x= path,replacement=paste("bayes/output/",mix,sep=''),fixed=T))
                        
          }
    }
   
   file.copy(from=paste0(dir,"/baseline/Kenai14pops96loci.bse"),to=paste0(dir,"/output/",mix,"/Kenai14pops96loci.bse"))
   
   file.copy(from=bayesexe.paths,to=paste0("bayes/output/",mix,"/",bayesexe.files))
   
 }  


```
###Summarize BAYES mixture results
```{r BAYES mixture summary by group, echo=FALSE}

      
BAYES_results<-CustomCombineBAYESOutput.GCL(groupvec=seq(3),groupnames=groups, maindir="BAYES/output", mixvec=Mixtures, prior="", ext="RGN", nchains=5, burn=0.5, alpha=0.1,PosteriorOutput=FALSE)

results_sum_df<-BAYES_results%>% 
  sapply(FUN = as_tibble,USE.NAMES = TRUE,simplify = FALSE) %>% 
  bind_rows(.id="mixture") %>% 
  mutate(group=rep(groups,length(Mixtures))) %>% 
  select(mixture,group,mean,'5%','95%',sd)

results_sum_df


```
###Checking compostion by population
```{r BAYES mixture summary by pop, echo=FALSE}

      
BAYES_results_pop<-CustomCombineBAYESOutput.GCL(groupvec=1:14,groupnames=Kenai14pops, maindir="bayes/output", mixvec=Mixtures, prior="", ext="BOT", nchains=5, burn=0.5, alpha=0.1,PosteriorOutput=FALSE)

pop_results_sum_df<-BAYES_results_pop%>% 
  sapply(FUN = as_tibble,USE.NAMES = TRUE,simplify = FALSE) %>% 
  bind_rows(.id="mixture") %>% 
  mutate(pop=rep(PopNames,length(Mixtures))) %>% 
  select(mixture,pop,mean,'5%','95%',sd)

pop_results_sum_df

```

###Analyze mixtures with rubias
```{r rubias analysis,echo=FALSE}

priors_rubias<-tibble(collection=Kenai14pops,pi_param=priors,minval=0.01) 

sapply(Mixtures,function(mix){
  
  run_mix<-mix.in %>% filter(collection==mix)
  

  
  run_rubias_mixture(reference=baseline, mixture=run_mix, group_names=groups, gen_start_col=5, method = "PB", 
                               alle_freq_prior = list(const_scaled = 1), pi_prior =priors_rubias , 
                               pi_init = NULL, reps = 25000, burn_in = 5000, pb_iter = 100, 
                               sample_int_Pi = 10, pi_prior_sum = 1, 
                               path = "rubias/output", seed = 56) 
}) 

```
###Summarize rubias results
The rubias mcmc and bias correced results are very similar to the BAYES results, but using BAYES estimates for remaining analyses
```{r rubias mixture summary,echo=FALSE}

rubias_results_bias_cor <- custom_combine_rubias_output(rubias_output = NULL, mixvec = Mixtures, group_names = groups, 
    group_names_new = NULL, groupvec = NULL, groupvec_new = NULL, 
    path = "rubias/output", alpha = 0.1, burn_in = 5000, bias_corr = TRUE, 
    threshold = 5e-07, plot_trace = FALSE)

rubias_results_mcmc <- custom_combine_rubias_output(rubias_output = NULL, mixvec = Mixtures, group_names = groups, 
    group_names_new = NULL, groupvec = NULL, groupvec_new = NULL, 
    path = "rubias/output", alpha = 0.1, burn_in = 5000, bias_corr = FALSE, 
    threshold = 5e-07, plot_trace = FALSE)

results_all_methods <- bind_rows(
  rubias_results_bias_cor %>% 
  mutate(method="rubias_PB",lo5CI=`5%`,hi95CI=`95%`) %>% 
  select(-`P=0`,-`5%`,-`95%`,-median),

rubias_results_mcmc %>% 
  mutate(method="rubias_MCMC",lo5CI=`5%`,hi95CI=`95%`)%>% 
   select(-`P=0`,-`5%`,-`95%`,-median),

results_sum_df %>% 
  mutate(mixture_collection=mixture,repunit=factor(group,levels=groups),method="BAYES",lo5CI=`5%`,hi95CI=`95%`)%>%
  select(-group,-mixture,-`5%`,-`95%`)

) %>% 
  mutate(mixture_collection=factor(mixture_collection,levels=Mixtures))

results_all_methods
```
###Plot results
```{r plot estimates to compare, fig.height=10, fig.width=8}

GrCol <- c("blue","red","green")

results_all_methods %>% 
  ggplot(aes(y=mean,x=repunit,fill=repunit))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.text.x=element_text(size=14), 
          strip.text.y=element_text(size=14), 
          axis.title.x=element_text(size=13),
          axis.title.y=element_text(size=13),
          legend.position="none")+
  geom_bar(stat="identity",position = "dodge") +
  geom_errorbar(aes(ymin = lo5CI, ymax = hi95CI, width = 0.3), position = "dodge")+
  scale_fill_manual(name = "Reporting Group", values = GrCol)+
  facet_grid(mixture_collection~method) +
  ylab("Proportion (90% CI)")+
  xlab("Reporting Group")
  
ggsave(filename="output/2018_RussianRiverRunTimingStockComps_AllMethods.pdf",device = "pdf")

```
###Plot Bayes results by pop
```{r plot populuation estimats, fig.height=8, fig.width=5}

pop_results_sum_df %>% 
  mutate(mixture=factor(mixture,levels=Mixtures),pop=factor(pop,levels=PopNames)) %>% 
  ggplot(aes(y=mean,x=pop,fill=pop))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.text.x=element_text(size=14), 
          strip.text.y=element_text(size=14), 
          axis.title.x=element_text(size=13),
          axis.title.y=element_text(size=13),
          legend.position="none")+
  geom_bar(stat="identity",position = "dodge") +
  geom_errorbar(aes(ymin = `5%`, ymax = `95%`, width = 0.3), position = "dodge")+
  facet_grid(rows=vars(mixture))+
  ylab("Proportion (90% CI)")+
  xlab("Population")
  
ggsave(filename="output/2018_RussianRiverRunTimingStockComps_BAYES_ByPopulation.pdf",device = "pdf")

```
##Analyze mixtures in rubias with finer-scale reporting groups
```{r fine scale reporting groups}

groupvec6 <- c(1,1,1,1,1,2,3,4,4,4,4,2,5,6)
groups6 <- c("UpperTribs","UpperMainstem","RussianEarly","RussianLate","HiddenL","SkilakOutlet")

attach("V:/Analysis/2_Central/Sockeye/Cook Inlet/2012 Baseline/Baseline/EarlyRussianProoftests.RData")

baseline_6groups<-create_rubias_baseline(sillyvec = Kenai14pops,loci=loci93,group_names = groups6,groupvec = groupvec6,baseline_name = "Kenai14pops96loci6groups")#rubias

detach()
baseline_6groups

```
```{r rubias analysis,echo=FALSE}

sapply(Mixtures,function(mix){
  
  run_mix<-mix.in %>% filter(collection==mix)
  
  run_rubias_mixture(reference=baseline_6groups, mixture=run_mix, group_names=groups6, gen_start_col=5, method = "PB", 
                               alle_freq_prior = list(const_scaled = 1), pi_prior =priors_rubias , 
                               pi_init = NULL, reps = 25000, burn_in = 5000, pb_iter = 100, 
                               sample_int_Pi = 10, pi_prior_sum = 1, 
                               path = "rubias/output/6_groups", seed = 56) 
}) 

```
###Summarize rubias results
The rubias mcmc and bias correced results are very similar to the BAYES results, but using BAYES estimates for remaining analyses
```{r rubias mixture summary,echo=FALSE}

rubias_results_bias_cor_6groups <- custom_combine_rubias_output(rubias_output = NULL, mixvec = Mixtures, group_names = groups6, 
    group_names_new = NULL, groupvec = NULL, groupvec_new = NULL, 
    path = "rubias/output/6_groups", alpha = 0.1, burn_in = 5000, bias_corr = TRUE, 
    threshold = 5e-07, plot_trace = FALSE)

rubias_results_mcmc_6groups <- custom_combine_rubias_output(rubias_output = NULL, mixvec = Mixtures, group_names = groups6, 
    group_names_new = NULL, groupvec = NULL, groupvec_new = NULL, 
    path = "rubias/output/6_groups", alpha = 0.1, burn_in = 5000, bias_corr = FALSE, 
    threshold = 5e-07, plot_trace = FALSE)

results_all_methods_6groups <- bind_rows(
  
  rubias_results_bias_cor_6groups %>% 
  mutate(method="rubias_PB",lo5CI=`5%`,hi95CI=`95%`) %>% 
  select(-`P=0`,-`5%`,-`95%`,-median),

rubias_results_mcmc_6groups %>% 
  mutate(method="rubias_MCMC",lo5CI=`5%`,hi95CI=`95%`)%>% 
   select(-`P=0`,-`5%`,-`95%`,-median)

) %>% 
  mutate(mixture_collection=factor(mixture_collection,levels=Mixtures))

results_all_methods_6groups
```
###Plot results for 6 groups
```{r plot estimates to compare, fig.height=10, fig.width=8}

GrCol6 <- c("magenta","cyan","red","green","orange","grey")

results_all_methods_6groups %>% 
  ggplot(aes(y=mean,x=repunit,fill=repunit))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.text.x=element_text(size=14), 
          strip.text.y=element_text(size=14), 
          axis.title.x=element_text(size=13),
          axis.title.y=element_text(size=13),
          legend.position="none")+
  geom_bar(stat="identity",position = "dodge") +
  geom_errorbar(aes(ymin = lo5CI, ymax = hi95CI, width = 0.3), position = "dodge")+
  scale_fill_manual(name = "Reporting Group", values = GrCol6)+
  facet_grid(mixture_collection~method) +
  ylab("Proportion (90% CI)")+
  xlab("Reporting Group")
  
ggsave(filename="output/2018_RussianRiverRunTimingStockComps_AllMethods_6groups.pdf",device = "pdf")

```
###Plot rubias MCMC results for 3 and 6 groups
```{r plot estimates to compare, fig.height=10, fig.width=8}

GrCol_all <- c("red","green","blue","magenta","cyan","orange","grey")

levels7<-c("RussianEarly","RussianLate","KenaiOther","UpperTribs","UpperMainstem","HiddenL","SkilakOutlet")

df <-bind_rows( results_all_methods %>% 
                  filter(method=="rubias_MCMC") %>% 
                  mutate(n_groups="3 groups"),
                
                results_all_methods_6groups %>% 
                  filter(method=="rubias_MCMC") %>% 
                  mutate(n_groups="6 groups") 
) %>% 
  group_by(mixture_collection,n_groups) %>% 
  arrange(desc(n_groups),match(mixture_collection,Mixtures),match(repunit,levels7)) %>% 
  ungroup() %>% 
  mutate(repunit=factor(repunit,levels=levels7))
  

  df %>% 
  ggplot(aes(y=mean,x=repunit,fill=repunit))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.text.x=element_text(size=14), 
          strip.text.y=element_text(size=14), 
          axis.title.x=element_text(size=13),
          axis.title.y=element_text(size=13),
          legend.position="none")+
  geom_bar(stat="identity",position = "dodge") +
  geom_errorbar(aes(ymin = lo5CI, ymax = hi95CI, width = 0.3), position = "dodge")+
  scale_fill_manual(name = "Reporting Group", values = GrCol_all)+
  facet_grid(mixture_collection~n_groups,scales = "free_x",space = "free")+
  ylab("Proportion (90% CI)")+
  xlab("Reporting Group")
  
ggsave(filename="output/2018_RussianRiverRunTimingStockComps_rubias_MCMC_3vs6_groups.pdf",device = "pdf")

```